Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-27 paper.qmd
Added long edge plots

hex_grid_coloured_two_curvy2 <- ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts_s_curve2, 
    aes(x = x, y = y, 
        group = hex_poly_id, 
        fill = std_counts), 
        color = "grey70", 
        linewidth=0.2) +
  geom_point(data = umap_two_curvy_scaled,
           aes(x = UMAP1, y = UMAP2),
           alpha = 0.3,
           size = 0.5) +
  scale_fill_viridis_c(direction = -1, 
    na.value = "#ffffff", option = "C") +
  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
  interior_annotation("b")
```

@@

true_pred_df <- predict_emb(
  test_data = true_model_df,
  df_bin_centroids = df_bin_centroids_two_curvy,
  df_bin = df_bin_two_curvy,
  type_NLDR = "UMAP"
)

# Compute radius r
a1 <- calc_bins_y(
  bin1 = 15, 
  r2 = r2, 
  q = 0.1)$a1

@@

>>>>

hex_grid_coloured_two_curvy2 <- ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts_s_curve2, 
    aes(x = x, y = y, 
        group = hex_poly_id, 
        fill = std_counts), 
        color = "grey70", 
        linewidth=0.2) +
  geom_point(data = umap_two_curvy_scaled,
           aes(x = UMAP1, y = UMAP2),
           alpha = 0.3,
           size = 0.5) +
  scale_fill_viridis_c(direction = -1, 
    na.value = "#ffffff", option = "C") +
  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
  interior_annotation("b")

## Triangulate bin centroids
tr1_object_two_curvy2 <- tri_bin_centroids(
  df_bin_centroids2, x = "c_x", y = "c_y")
tr_from_to_df_two_curvy2 <- gen_edges(
  tri_object = tr1_object_two_curvy2)

trimesh_two_curvy2 <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy2,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#33a02c") +
  interior_annotation("a", position = c(0.92, 0.95))

## Compute 2D distances
distance_two_curvy <- cal_2d_dist(
  tr_coord_df = tr_from_to_df_two_curvy2, 
  start_x = "x_from", 
  start_y = "y_from", 
  end_x = "x_to", 
  end_y = "y_to", 
  select_vars = c("from", "to", "distance"))

## Benchmark 1

distance_df_small_edges_two_curvy2 <- distance_two_curvy |>
  filter(distance < benchmark_two_curvy) ## 0.231

tr_from_to_df_two_curvy3 <- inner_join(
  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy2,
  by = c("from", "to"))

trimesh_two_curvy_removed1 <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy3,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#33a02c") +
  interior_annotation("c", position = c(0.92, 0.95))

## Benchmark 2

benchmark_two_curvy <- 1.3

distance_df_small_edges_two_curvy2 <- distance_two_curvy |>
  filter(distance < benchmark_two_curvy)

tr_from_to_df_two_curvy4 <- inner_join(
  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy2,
  by = c("from", "to"))

trimesh_two_curvy_removed2 <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy4,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#33a02c") +
  interior_annotation("b", position = c(0.92, 0.95))

@@

true_pred_df <- predict_emb(
  test_data = true_model_df,
  df_bin_centroids = df_bin_centroids2,
  df_bin = df_bin_two_curvy2,
  type_NLDR = "UMAP"
)

# Compute radius r
a1 <- calc_bins_y(
  bin1 = 13, 
  r2 = r2, 
  q = 0.1)$a1

@@

```{r}
#| fig-width: 6
#| fig-height: 4
#| fig-cap: "long edge removal"

trimesh_two_curvy2 + trimesh_two_curvy_removed2 + trimesh_two_curvy_removed1 +
  plot_layout(ncol = 3)
```
